---
title: "WHD_B_PreProcessing"
output: html_notebook
author: 'Francesca Barbieri'
---

```{r setup, include = FALSE, print = FALSE}

require("knitr")
opts_knit$set(root.dir = "C:/Users/Francesca/Documents/INTERNSHIP/") 

```

```{r, echo = FALSE, include = FALSE, print = FALSE}

library(readxl)
library(dplyr)
library(purrr)
library(tidyverse)
library(reshape2) # melt function
library(plyr)

library(ggplot2)
library(ggfortify)

library(pals)

# Ecology package
library(vegan)
library(phyloseq)
library(microbiome)

```

```{r, print = FALSE, include=FALSE}

bact_init_classification <- read_excel('Data/WHD_data_files/WHD_data_files/WHD_Bv6_paired_ASV_run_4MSc/asv/taxonomy_dada2/taxonomy_assignment_noSingletons.xlsx', col_names = FALSE, na = "NA")

```

# OTU table

```{r}

# From the file.txt, the # has been deleted to allow the header names
asv_table_bact <- read.table('Data/WHD_data_files/WHD_data_files/WHD_Bv6_paired_ASV_run_4MSc/asv/taxonomy_dada2/asvTable_noSingletons.txt', sep = '\t', header = TRUE)

#asv_table_bact$OTU.ID <- paste0('b_', asv_table_bact$OTU.ID)
rownames(asv_table_bact) <- asv_table_bact$OTU.ID

asv_table_bact <- asv_table_bact[, !(names(asv_table_bact) %in% c("taxonomy", "OTU.ID"))]

asv_table_bact <- t(asv_table_bact)
# Rows = Samples
# Columns = OTU

# 25689 51

```

# Classification

```{r, print = FALSE, include=FALSE}

classification <- read_excel('Data/WHD_data_files/WHD_data_files/WHD_Bv6_paired_ASV_run_4MSc/asv/taxonomy_dada2/taxonomy_assignment_noSingletons.xlsx', col_names = FALSE, na = "NA")
# TIBBLE 25689 x 26

```

```{r}

classification <- classification[,!names(classification) %in% c('...2','...3','...4','...5', '...8', '...11', '...23', '...26')]

# Checking entries
# all(classification$...7 == 'domain') # TRUE
# all(classification$...10 == 'phylum') # TRUE

colnames(classification)[colnames(classification) == '...1'] <- 'OTU'
colnames(classification)[colnames(classification) == '...6'] <- 'domain'
colnames(classification)[colnames(classification) == '...9'] <- 'phylum'
colnames(classification)[colnames(classification) == '...12'] <- 'class'
colnames(classification)[colnames(classification) == '...15'] <- 'order'
colnames(classification)[colnames(classification) == '...18'] <- 'family'

# Remove chloroplasts
chloroplasts <- subset(classification, grepl("Chloroplast", classification$class, ignore.case = TRUE))


classification <- subset(classification, !grepl("Chloroplast", classification$class, ignore.case = TRUE))

classification <- classification[,!names(classification) %in% c('...7', '...14', '...17', '...20', '...11')]

head(classification)

```

## Chloroplasts

```{r}

#unique(chloroplasts$phylum) # Cyanobacteriota
#unique(chloroplasts$class)  # Chloroplast
#unique(chloroplasts$order)  # Chloroplast
#unique(chloroplasts$family) # Bacillariophyta, Cryptomonadaceae, Euglenida,                                 Chlorarachniophyceae, Chlorophyta, Streptophyta,                               Bangiophyceae


colnames(chloroplasts)[colnames(chloroplasts) == 'family'] <- 'genus'
colnames(chloroplasts)[colnames(chloroplasts) == '...14'] <- 'class %'
colnames(chloroplasts)[colnames(chloroplasts) == '...17'] <- 'fam %'
colnames(chloroplasts)[colnames(chloroplasts) == '...20'] <- 'genus %'
colnames(chloroplasts)[colnames(chloroplasts) == 'order'] <- 'family' 

chloroplasts <- chloroplasts[,!names(chloroplasts) %in% c('...7', '...10', '...13', '...16', '...19', '...21', '...22', '...24', '...25')]

head(chloroplasts)

```

## Archaea and eukaryotes

Eukaryotes and archaea should be drop later anyway

```{r}

#unique(classification$domain) # FALSE

archaea <- classification[grep('archaea', classification$domain, ignore.case = TRUE), ]

eukaryota <- classification[grep('eukaryota', classification$domain, ignore.case = TRUE),]

classification <- subset(classification, !grepl('archaea', classification$domain, ignore.case = TRUE))
classification <- subset(classification, !grepl('eukaryota', classification$domain, ignore.case = TRUE))

# unique(classification$domain) # Bacteria

head(archaea)
head(eukaryota)
```

## Main DF

Matching 'unphased' entries.

```{r}
#unique(classification$...10) # phylum
#unique(classification$...13) # class, genus, family

to_match <- classification[classification$...13 != 'class',]

classification <- classification[classification[['...13']] == 'class', ]
#unique(classification$...13) # class

#unique(classification$...16) # order, genus, family

to_match <- classification[classification$...16 != 'order',]

classification <- classification[classification$...16 == 'order', ]
#unique(classification$...16) # order

#unique(classification$...19) # family, genus, suborder

to_match <- classification[classification$...19 != 'family',]

classification <- classification[classification$...19 == 'family', ]
#unique(classification$...19) # family

#unique(classification$...22) #genus

colnames(classification)[colnames(classification) == '...21'] <- 'genus'

classification <- classification[,!names(classification) %in% c('...10', '...13','...16','...19','...22', '...24','...25','...26')]

head(classification)
# 23489 x 7 
```

```{r}

# From family on
#unique(to_match$...19) # genus, suborder

#
to_match_genus <- (split(to_match, to_match$...19))$genus

#unique(to_match_genus$...19)  #genus
#unique(to_match_genus$...22) # NA
#unique(to_match_genus$...25) # NA

colnames(to_match_genus)[colnames(to_match_genus) == 'family'] <- 'genus'
to_match_genus <- to_match_genus[!(names(to_match_genus) %in% paste0("...", 19:26))]
to_match_genus <- to_match_genus[,!names(to_match_genus) %in% c('...10','...13','...16')]

taxonomy_classification_bact <- bind_rows(classification, to_match_genus)

#
to_match_suborder <- (split(to_match, to_match$...19))$suborder
colnames(to_match_suborder)[colnames(to_match_suborder) == 'family'] <- 'suborder'
colnames(to_match_suborder)[colnames(to_match_suborder) == '...21'] <- 'family'
colnames(to_match_suborder)[colnames(to_match_suborder) == '...24'] <- 'genus'

#unique(to_match_suborder$...19) # suborder
#unique(to_match_suborder$...22) # family
#unique(to_match_suborder$...25) # genus

to_match_suborder <- to_match_suborder[,!names(to_match_suborder) %in% c('...10','...13','...16', '...19', '...22','...25')]

taxonomy_classification_bact <- as.data.frame(bind_rows(taxonomy_classification_bact, to_match_suborder))
taxonomy_classification_bact <- taxonomy_classification_bact[,!names(taxonomy_classification_bact) %in% c('suborder','...3','...5','...7','...9','...10','...11', '...13', '...15', '...4', '...6', '...8', '...12', '...14')]


#prefix_b <- paste0('b_', taxonomy_classification_bact[,1])
#row.names(taxonomy_classification_bact) <- prefix_b
rownames(taxonomy_classification_bact) <- taxonomy_classification_bact[,1]
taxonomy_classification_bact <- taxonomy_classification_bact[,-1]

head(taxonomy_classification_bact)
# 24461 OTU
# 6 classification levels

write.table(taxonomy_classification_bact, 'Data/bact.classification.txt')

```

# Metadata

```{r}

metadata <- read_excel('Data/WHD_data_files/WHD_data_files/WHD_MIMARKS_Table/WHD_Bv6_Ev4_2013_Metadata_treatment.xlsx',
                       na = "NA")

metadata <- as.data.frame(t(metadata))

colnames(metadata) <- c('project_name', 'investigation_type', 'experimental_factor', 'geo_loc_name', 'lat_lon', 'collection_date','biome', 'feature', 'material', 'env_package', 'samp_collect_device', 'samp_size', 'nucl_acid_ext', 'lib_reads_seqd', 'lib_const_meth', 'target_gene', 'target_subfragment', 'seq_meth', 'pcr_primers', 'pcr_cond', 'seq_quality_check', 'sop', 'url', 'depth (m)', 'temp (C)', 'salinity (ppt)')

rownames(metadata) <- substr(rownames(metadata), 1, 8)


metadata$collection_date <- as.Date(as.numeric(metadata$collection_date), origin = "1899-12-30")

metadata$collection_date <- as.factor(metadata$collection_date)
```

## Sample checking

```{r}

bact_sample.depth <- rowSums(asv_table_bact)
bact_sample.depth <- bact_sample.depth[order(bact_sample.depth)]

#summary(bact_sample.depth)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#  7431   46698   78266  225772  153531 2746325 

# Cut off at 20.000 (?)

b.metadata <- data.frame(sample = rownames(metadata), 
                         collection_date = metadata$collection_date,
                         sample_depth = bact_sample.depth,
                         material = metadata$material)

b.metadata <- b.metadata %>%
  group_by(collection_date, material) %>%
  mutate(replicates = dense_rank(paste(collection_date, material)))

# They remain consistent per biological replicate

```

# Phyloseq object

```{r}

OTU_b = otu_table(as.matrix(asv_table_bact), taxa_are_rows = FALSE)
SAM = sample_data(metadata) 
TAX_b = tax_table(as.matrix(taxonomy_classification_bact))

bact_phylo <- phyloseq(OTU_b, TAX_b, SAM)

bact_phylo
# 24461 taxa x 51 samples

```

# PRE-PROCESSING

```{r}

# Normalization
bact_n <- transform_sample_counts(bact_phylo, function(x) x / sum(x))

# Sequencing error filtering
bact_nf <- filter_taxa(bact_n, function(x) {sum(x > 0) > 1 & max(x) > 0.01}, TRUE) 
# 109

# Normalization 
bact_nfn <- transform_sample_counts(bact_nf, function(x) x / sum(x))

# CLR
bact_nfc <- microbiome::transform(bact_nf, "clr")

```

Using standard statistical methods on such data may lead to spurious results. Likewise, any data that are compositional in nature are interpreted into relative counts. Hence, using a CLR transformation allows the circumvention of these spurious results 

```{r}

taxa_names <- colnames(as.data.frame(bact_nf@otu_table))  # or use colnames() if it's in columns
bact_f <- prune_taxa(taxa_names, bact_phylo)

```


## Rarefaction

```{r, eval = FALSE}

b_otu <- as.data.frame(bact_phylo@otu_table)

b_min_read_sample <- min(rowSums(b_otu))
b_min_sample_index <- which(rowSums(b_otu) == b_min_read_sample)
#WHD_00070, 7431

b_colors <- rep("lightblue", nrow(b_otu))
b_colors[b_min_sample_index] <- "blue"

#pdf('Scripts/Bacteria/Res/B_PreProcessing/B_rarefaction.pdf',10, 10)
bact_rarecurve <- rarecurve(b_otu, step = 500, label = FALSE, cex = 0.75, las = 1, col = b_colors)
abline(v = b_min_read_sample, lty = 'dashed', col = 'black')
#dev.off()

```

```{r, eval = FALSE}

min_read_sample <- min(rowSums(asv_table_bact))
#WHD_00070, 7431

s <- specnumber(asv_table_bact)
bact_rarefied <- rarefy(asv_table_bact, min_read_sample)

#pdf('Scripts/Bacteria/Res/B_PreProcessing/B_rarefied.species.pdf')
plot(s, bact_rarefied, xlab = "Observed # Species", ylab = "Rarefied # Species", main = 'Rarefied species - Bacteria')
abline(0,1)
#dev.off()

```


# Graphs

## PCA

```{r}

# CLR!

otu <- as.data.frame(bact_nfc@otu_table)

pca <- prcomp(otu)#, scale. = TRUE)

otu$material <- metadata$material
otu$collection_date <- metadata$collection_date

autoplot(pca, data = otu, colour = 'collection_date', shape = 'material') +
  ggtitle('PCA - Bacteria') +
  theme_minimal()

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCA.pdf')

```


## Compositionality

```{r, eval = FALSE}

library(Polychrome)

P40 <- createPalette(40, c("#a2da5a","#a877c8", "#a2b2df"), range = c(30, 80))
P40 <- sortByHue(P40)
P40 <- as.vector(t(matrix(P40, ncol=4)))

bact_m <- psmelt(bact_nfn)
material_lab <- setNames(bact_m$material, bact_m$Sample)

ggplot(bact_m, aes(x = Sample, y = Abundance, fill = phylum)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Material", y = "Relative Abundance\n", fill = "Phylum") +
  scale_fill_manual(values = P40) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 18),
    plot.title = element_text(size = 20, hjust = 0.5, family = "sans")
  ) +
  scale_x_discrete(labels = material_lab) +
  facet_wrap(~ collection_date, scales = "free_x", nrow = 1)

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_composition.pdf', height = 8, width = 15)

```

#### Pseudomonadota

```{r}
proteobacteria <- subset_taxa(bact_nfn, phylum == "Pseudomonadota")
#83

# Roseobacteraceae   22
# Paracoccaceae   12
# Halieaceae    4
# Hyphomonadaceae    3

taxonomy <- as.data.frame(tax_table(proteobacteria))

taxonomy$family <- ifelse(
  taxonomy$family %in% c("Roseobacteraceae", "Paracoccaceae", 'Hyphomonadaceae','Halieaceae'),  taxonomy$family,  # Keep these families
  "Others")

# Assign back to the phyloseq object
tax_table(proteobacteria) <- as.matrix(taxonomy)


```


```{r}
proteobacteria_rel <- transform_sample_counts(proteobacteria, function(x) x / sum(x))

proteobacteria_family <- tax_glom(proteobacteria_rel, taxrank = "family")
proteobacteria_family <- psmelt(proteobacteria_family)

```

```{r}

P40 <- c('#C36B85', '#FB938F', '#F2CACB', '#FDBB75', 'lightgrey')

ggplot(proteobacteria_family, aes(y = Sample, x = Abundance, fill = factor(family, levels = c("Roseobacteraceae", "Paracoccaceae", 'Hyphomonadaceae','Halieaceae', "Others")))) +
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_manual(
    values = P40,
    breaks = c("Roseobacteraceae", "Paracoccaceae", 'Hyphomonadaceae','Halieaceae', "Others"),
    drop = FALSE # Ensure all levels, even unrepresented ones, appear in the legend
  ) +
  scale_x_continuous(labels = scales::percent) +
  labs(y = "Sample", x = "Relative Abundance (%)", fill = "Family") +
  theme_bw()+
   theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8, angle = 0),
    legend.position = 'none'
  ) +
  scale_y_discrete(labels = material_lab) +
  facet_wrap(~ collection_date, scales = "free_y", ncol = 1)


ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_proteobacteria.pdf', height = 15, width = 5)

```



## Richness

```{r}

plot_richness(bact_f, "material", measures= 'Shannon', color = 'material') +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) + 
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +  
  stat_summary(fun = mean, geom = "point", color = "black", size = 1.5) +  # mean points
  facet_wrap(~ collection_date, scales = 'fixed', nrow = 1) +
  scale_color_brewer(palette = "Set1") +  
  labs(y = "Shannon Index", x = NULL) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom",
    axis.text.x = element_blank()
  )

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_Shannon.richness.pdf', width = 40, height = 20, units = "cm")

```

```{r}

plot_richness(bact_f, "material", measures= 'Observed', color = 'material') +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) + 
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +  
  stat_summary(fun = mean, geom = "point", color = "black", size = 1.5) +  # mean points
  facet_wrap(~ collection_date, scales = 'fixed', nrow = 1) +
  scale_color_brewer(palette = "Set1") +  
  labs(y = "Observed Richness", x = NULL) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom",
    axis.text.x = element_blank()
  )

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_observed.richness.pdf', width = 60, height = 30, units = "cm")

```


```{r}

# Chao2

library(fossil)

otu <- as.data.frame(bact_f@otu_table)

# Chao2 per sample
c2 <- apply(otu, 1, chao2)

sample_data(bact_f)$chao2 <- c2

ggplot(sample_data(bact_f), aes(x = material, y = chao2, color = material)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) + 
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +  
  stat_summary(fun = mean, geom = "point", color = "black", size = 1.5) +  # mean points
  facet_wrap(~ collection_date, scales = 'fixed', nrow = 1) +
  scale_color_brewer(palette = "Set1") +  
  labs(y = "Chao2 Index", x = NULL) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12),
    legend.position = "bottom",
    axis.text.x = element_blank()
  )

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_Chao2.richness.pdf', width = 40, height = 20, units = "cm")

```



## PCoA

```{r}

# Jaccard
pcoa <- ordinate(bact_nfn, method = "PCoA", distance = "jaccard")
plot_ordination(bact_nfn, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCoA1 [22.7%]') +
  #ylab('PCoA2 [13.9%]') +
  ggtitle('Pricipal Coordinate Analysis - Jaccard')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_jcrd.pdf')

# Horn
pcoa <- ordinate(bact_nfn, method = "PCoA", distance = "horn")
plot_ordination(bact_nfn, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCoA1 [41.5%]') +
  #ylab('PCoA2 [23.8%]') +
  ggtitle('Pricipal Coordinate Analysis - Horn')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_horn.pdf')

# Morisita
pcoa <- ordinate(bact_nfn, method = "PCoA", distance = "morisita")
plot_ordination(bact_nfn, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCoA1 [2%]') +
  #ylab('PCoA2 [2%]') +
  ggtitle('Pricipal Coordinate Analysis - Morisita')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_morisita.pdf')

# Bray-Curtis
pcoa <- ordinate(bact_nfn, method = "PCoA", distance = "bray")
plot_ordination(bact_nfn, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette = "RdBu") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5, family = "sans")
  ) +
  xlab('PCo1 [32.7%]') +
  ylab('PCo2 [16.8%]') +
  ggtitle('PCoA - Bray-Curtis')
  
#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_bray.pdf')


```

### No normalization

```{r}

# Jaccard
pcoa <- ordinate(bact_phylo, method = "PCoA", distance = "jaccard")
plot_ordination(bact_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCoA1 [18.7%]') +
  #ylab('PCoA2 [12.7%]') +
  ggtitle('Pricipal Coordinate Analysis - Jaccard (Not normalizad)')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_jcrd_NOTnorm.pdf')


# Bray-Curtis
pcoa <- ordinate(bact_phylo, method = "PCoA", distance = "bray")
plot_ordination(bact_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette = "RdBu") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5, family = "sans")
  ) +
  #xlab('PCo1 [29%]') +
  #ylab('PCo2 [15.9%]') +
  ggtitle('PCoA - Bray-Curtis')
  
#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_bray_NOTnorm.pdf')


# Horn
pcoa <- ordinate(bact_phylo, method = "PCoA", distance = "horn")
plot_ordination(bact_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCoA1 [37%]') +
  #ylab('PCoA2 [21.6%]') +
  ggtitle('Pricipal Coordinate Analysis - Horn (Not normalizad)')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_horn_NOTnorm.pdf')


# Morisita
pcoa <- ordinate(bact_phylo, method = "PCoA", distance = "morisita")
plot_ordination(bact_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCoA1 [37.1%]') +
  #ylab('PCoA2 [21.1%]') +
  ggtitle('Pricipal Coordinate Analysis - Morisita(Not normalizad)')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_PCoA_morisita_NOTnorm.pdf')

```

## NMDS

```{r}

# Jaccard
nmds <- ordinate(bact_nfn, method = "NMDS", distance = "jaccard")
saveRDS(nmds, 'Scripts/Bacteria/Res/B_PreProcessing/NMDS_stress/B_NMDS.jcrd.rds')
plot_ordination(bact_nfn, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Jaccard')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_NMDS.jcrd.pdf')


# Horn
nmds <- ordinate(bact_nfn, method = "NMDS", distance = "horn")
saveRDS(nmds, 'Scripts/Bacteria/Res/B_PreProcessing/NMDS_stress/B_NMDS.horn.rds')
plot_ordination(bact_nfn, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Horn')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_NMDS.horn.pdf')


# Bray-Curtis

nmds <- ordinate(bact_nfn, method = "NMDS", distance = "bray")
saveRDS(nmds, 'Scripts/Bacteria/Res/B_PreProcessing/NMDS_stress/B_NMDS.bray.rds')
plot_ordination(bact_nfn, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette = "RdBu") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5, family = "sans")
  ) +
  #facet_wrap(~ collection_date, scales = 'free_x') +
  annotate("text", x = -Inf, y = Inf, label = "Stress = 0.118", hjust = -0.1, vjust = 1.5, size = 5, color = "black")
  ggtitle('Bacteria NMDS')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_NMDS.bray.pdf',height = 7, width = 9)
  
```

### No normalization

```{r}

# Jaccard
nmds <- ordinate(bact_phylo, method = "NMDS", distance = "jaccard")
saveRDS(nmds, 'Scripts/Bacteria/Res/B_PreProcessing/NMDS_stress/B_NMDS.jcrd_NOTnorm.rds')
plot_ordination(bact_phylo, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Jaccard (Not normalizad)')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_NMDS_NOTnorm.jcrd.pdf')


# Horn
nmds <- ordinate(bact_phylo, method = "NMDS", distance = "horn")
saveRDS(nmds, 'Scripts/Bacteria/Res/B_PreProcessing/NMDS_stress/B_NMDS_NOTnorm.horn.rds')
plot_ordination(bact_phylo, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Horn (Not normalizad)')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_NMDS.horn_NOTnorm.pdf')


# Bray-Curtis

nmds <- ordinate(bact_phylo, method = "NMDS", distance = "bray")
saveRDS(nmds, 'Scripts/Bacteria/Res/B_PreProcessing/NMDS_stress/B_NMDS_NOTnorm.bray.rds')
plot_ordination(bact_phylo, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Bray-Curtis (Not normalizad)')

#ggsave('Scripts/Bacteria/Res/B_PreProcessing/B_NMDS.bray_NOTnorm.pdf')
  

```

# Saves

```{r}

# PHYLOSEQ OBJECTS
saveRDS(bact_phylo, 'Scripts/Bacteria/Res/B_PreProcessing/bact_phylo.rds')
saveRDS(bact_n, 'Scripts/Bacteria/Res/B_PreProcessing/bact_n.rds')
saveRDS(bact_nf, 'Scripts/Bacteria/Res/B_PreProcessing/bact_nf.rds')
saveRDS(bact_nfn, 'Scripts/Bacteria/Res/B_PreProcessing/bact_nfn.rds')

```








