---
title: "WHD_E_PreProcessing"
output: html_notebook
---

```{r setup, include = FALSE, print = FALSE}

require("knitr")
opts_knit$set(root.dir = "C:/Users/Francesca/Documents/INTERNSHIP/") 

```

```{r, echo = FALSE, include = FALSE, print = FALSE}

library(readxl)
library(dplyr)
library(purrr)
library(tidyverse)
library(reshape2) # melt function

library(ggplot2)
library(ggfortify)
library(pals)

# Ecology package
library(vegan)
library(phyloseq)
library(microbiome)
library(microViz)
library(fossil)

```

# Import

```{r}

euk_phylo <- readRDS('Scripts/Eukaryotes/Res/E_PreProcessing/euk_phylo.rds')
bact_phylo <- readRDS('Scripts/Bacteria/Res/B_PreProcessing/bact_phylo.rds')

euk_nfn <- readRDS('Scripts/Eukaryotes/Res/E_PreProcessing/euk_nfn.rds')
bact_nfn <- readRDS('Scripts/Bacteria/Res/B_PreProcessing/bact_nfn.rds')

#e.agg <- readRDS('Scripts/Eukaryotes/Res/E_SPIEC-EASI/euk_agglomerated.2.rds')
# normalized, filterd, normalized but og dataframe

```

# Phyloseq

```{r}

# Initial dataframe, containing all OTUs

# euk = 6 862
# bact = 24 461

ps_phylo <- merge_phyloseq(euk_phylo, bact_phylo)

ps_phylo
# 31 321

```

```{r}

# Dataframe after 0.01% filtering and normalized

# euk = 109
# bact= 130

ps <- merge_phyloseq(euk_nfn, bact_nfn)

ps # 239

```

## Rarefaction

```{r, eval = FALSE}

otu <- as.data.frame(ps_phylo@otu_table)

min_read_sample <- min(rowSums(otu))
#WHD_00070, 7431

#pdf('Scripts/  

rrcurve <- rarecurve(otu, step = 500, label = FALSE, cex = 0.75, las = 1, col = 'blue')
title('Rarefaction Curve - Bacteria')
#dev.off()

```

## Environmental Variables

```{r}

metadata <- read_excel('Data/WHD_data_files/WHD_data_files/WHD_MIMARKS_Table/WHD_Bv6_Ev4_2013_Metadata_treatment.xlsx',
                       na = "NA")

metadata <- as.data.frame(t(metadata))

colnames(metadata) <- c('project_name', 'investigation_type', 'experimental_factor', 'geo_loc_name', 'lat_lon', 'collection_date','biome', 'feature', 'material', 'env_package', 'samp_collect_device', 'samp_size', 'nucl_acid_ext', 'lib_reads_seqd', 'lib_const_meth', 'target_gene', 'target_subfragment', 'seq_meth', 'pcr_primers', 'pcr_cond', 'seq_quality_check', 'sop', 'url', 'depth', 'temp', 'salinity')

rownames(metadata) <- substr(rownames(metadata), 1, 8)


metadata$collection_date <- as.Date(as.numeric(metadata$collection_date), origin = "1899-12-30")

metadata$collection_date <- as.factor(metadata$collection_date)

```

```{r}

col.to.keep <- c('temp', 'salinity', 'collection_date', 'material')

md <- metadata[,col.to.keep]

md$temp <- as.numeric(md$temp)
md$salinity <- as.numeric(md$salinity)

```

```{r}

md.plot <- md %>%
  pivot_longer(cols = c('salinity', 'temp'),
               names_to = "Variable",
               values_to = "Value")

ggplot(md, aes(x = collection_date, y = temp)) + 
  geom_line(color = "blue", size = 1) +   # Add a blue line for temperature
  geom_point(color = "red", size = 2) +   # Add red points to the line
  labs(title = "Temperature Over Collection Dates",
       x = "Collection Date", 
       y = "Temperature (°C)") + 
  theme_minimal()

#ggsave('Scripts/temperature.pdf')
```

# Graphs

## PCA

Using standard statistical methods on such data may lead to spurious results. Likewise, any data that are compositional in nature are interpreted into relative counts. Hence, using a CLR transformation allows the circumvention of these spurious results 

```{r}

ps_c <- microbiome::transform(ps, "clr")

```

```{r}

otu <- as.data.frame(ps_c@otu_table)

pca <- prcomp(otu)#, scale. = TRUE)

otu$material <- metadata$material
otu$collection_date <- metadata$collection_date

# Plot the PCA results
autoplot(pca, data = otu, colour = 'collection_date', shape = 'material', size = 4) +
  ggtitle('PCA') +
  scale_color_brewer(palette = 'Set1') +
  theme(legend.position = 'bottom') +
  theme_minimal()

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE_PCA.pdf', height = 7, width = 9)

```

## Compositionality

```{r}
 
# COMPOSITIONALITY

plot_bar(ps, fill = 'phylum') +
  geom_col(aes(color=domain),stat = "identity", position = "stack") +
  labs(x = "", y = "Relative Abundance\n") +
  scale_fill_manual(values=as.vector(polychrome())) +
  ggtitle('Compositionality')

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE_compositionality.pdf', width = 50, height = 30, units = "cm")


```

```{r}

physeq_subset <- subset_samples(ps, collection_date == "2013-07-17")

# Transform to relative abundances (compositionally)
physeq_rel <- transform_sample_counts(physeq_subset, function(x) x / sum(x))


plot_bar(physeq_rel, x = "Sample", fill = "phylum") +
  theme_minimal() +
   geom_col(aes(color=phylum),stat = "identity", position = "stack") +
  labs(title = "Compositional Bar Plot for Selected Collection Date")

```

## Richness

```{r}

otu <- as.data.frame(ps@otu_table)

# Present/absent matrix
pres.mat <- otu > 0
pres.mat <- as.numeric(pres.mat)
pres.mat <- matrix(pres.mat, nrow = nrow(otu), ncol = ncol(otu),        dimnames = dimnames(otu))

```

```{r, eval = FALSE}

plot_richness(ps_phylo, "material", measures= c('Chao1','Shannon', 'Chao2')) +
  geom_boxplot() + 
  facet_wrap(collection_date ~ variable, scale = 'free_y')

# Shannon = abundances and evenness

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/be_richness.pdf')
```

## PCoA

```{r}

# Jaccard
pcoa <- ordinate(ps, method = "PCoA", distance = "jaccard")
plot_ordination(ps, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set2') +
  #xlab('PCo1 [15%]') +
  #ylab('PCo2 [13.3%]') +
  ggtitle('Principal Coordinate Analysis (ps) - Jaccard')

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_jcrd.pdf')

# Bray - Curtis
pcoa <- ordinate(ps, method = "PCoA", distance = "bray")
plot_ordination(ps, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette='Set1') +
  xlab('PCo1 [21.6%]') +
  ylab('PCo2 [17.6%]') +
  ggtitle('PCoA - Bray-Curtis') +
  theme_minimal()

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_bray.pdf', height = 7, width = 9)


# Horn 
pcoa <- ordinate(ps, method = "PCoA", distance = "horn")
plot_ordination(ps, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette='Set1') +
  xlab('PCo1 [34.3%]') +
  ylab('PCo2 [19%]') +
  ggtitle('PCoA - Horn-Morisita') +
  theme_minimal()

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_horn.pdf', height = 7, width = 9)

# Morisita
pcoa <- ordinate(ps, method = "PCoA", distance = "morisita")
plot_ordination(ps, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set2') +
  #xlab('PCo1 [3.7%]') +
  #ylab('PCo2 [3.6%]') +
  ggtitle('PCoA - Morisita')

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_morisita.pdf')

```

### Not normalized

```{r}

# Jaccard
pcoa <- ordinate(ps_phylo, method = "PCoA", distance = "jaccard")
plot_ordination(ps_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCo1 [15%]') +
  #ylab('PCo2 [13.3%]') +
  ggtitle('Principal Coordinate Analysis (ps_phylo) - Jaccard')

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_jcrd_NOTnorm.pdf', height = 7, width = 9)

# Bray - Curtis
pcoa <- ordinate(ps_phylo, method = "PCoA", distance = "bray")
plot_ordination(ps_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCo1 [16.8%]') +
  #ylab('PCo2 [14.5%]') +
  ggtitle('Principal Coordinate Analysis (ps_phylo) - Bray-Curtis')

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_bray_NOTnorm.pdf', height = 7, width = 9)


# Horn 
pcoa <- ordinate(ps_phylo, method = "PCoA", distance = "horn")
plot_ordination(ps_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCo1 [29.6%]') +
  #ylab('PCo2 [26%]') +
  ggtitle('Principal Coordinate Analysis (ps_phylo) - Horn-Morisita')

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_horn_NOTnorm.pdf', height = 7, width = 9)

# Morisita
pcoa <- ordinate(ps_phylo, method = "PCoA", distance = "morisita")
plot_ordination(ps_phylo, pcoa, color="collection_date", shape = 'material') +
  geom_point(size = 3) +
  scale_color_brewer(palette='Set1') +
  #xlab('PCo1 [19.6%]') +
  #ylab('PCo2 [29.6%]') +
  ggtitle('Principal Coordinate Analysis (ps_phylo) - Morisita')

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_PCoA_horn_NOTnorm.pdf', height = 7, width = 9)

```

## NMDS

```{r}

# Jaccard
nmds <- ordinate(ps, method = "NMDS", distance = "jaccard")
saveRDS(nmds, 'Scripts/Bact.Euk/Res/BE_1.PreProcessing/NMDS_stress/BE_nmds_jcrd')
plot_ordination(ps, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Jaccard')+
   theme_minimal()

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_NMDS_jcrd.pdf')


# Horn
nmds <- ordinate(ps, method = "NMDS", distance = "horn")
saveRDS(nmds, 'Scripts/Bact.Euk/Res/BE_1.PreProcessing/NMDS_stress/BE_nmds_horn')
plot_ordination(ps, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Horn')+
   theme_minimal()

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_NMDS_horn.pdf')


# Bray-Curtis
nmds <- ordinate(ps, method = "NMDS", distance = "bray", k = 3)
saveRDS(nmds, 'Scripts/Bact.Euk/Res/BE_1.PreProcessing/NMDS_stress/BE_nmds_bray')
plot_ordination(ps, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette='Set1') +
  labs(title = 'Bacteria + Eukaryotes',
       subtitle = 'NMDS - Bray-Curtis') +
   theme_minimal()

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_NMDS_bray.pdf',height = 7, width = 9)


# Morisita
nmds <- ordinate(ps, method = "NMDS", distance = "morisita")
saveRDS(nmds, 'Scripts/Bact.Euk/Res/BE_1.PreProcessing/NMDS_stress/BE_nmds_morisita')
plot_ordination(ps, nmds, color="collection_date", shape = 'material') +
  geom_point(size = 4) +
  scale_color_brewer(palette='Set1') +
  ggtitle('NMDS - Morisita') +
   theme_minimal()

#ggsave('Scripts/Bact.Euk/Res/BE_1.PreProcessing/BE.ps_NMDS_morisita.pdf')
```

```{r}

dissimilarity_matrix <- distance(ps, method = "bray")
nmds_result <- metaMDS(dissimilarity_matrix, k = 3, trymax = 100)

nmds_scores <- as.data.frame(scores(nmds_result))
nmds_scores$Sample <- rownames(nmds_scores)  # Add sample names

sample_data_df <- as.data.frame(ps@sam_data)

# Merge NMDS scores with sample metadata
nmds_scores <- merge(nmds_scores, sample_data_df, by.x = "Sample", by.y = "row.names")


# Create a 3D NMDS plot
plot_ly(data = nmds_scores, 
        x = ~NMDS1, 
        y = ~NMDS2, 
        z = ~NMDS3, 
        text = ~Sample, 
        color = ~collection_date,  # Color by collection_date
              symbol = ~material,        # Shape by material
              colors = "Set1",           # Choose a color palette
              type = 'scatter3d', 
              mode = 'markers', 
              marker = list(size = 5)) %>%
  layout(title = "3D NMDS Plot",
         scene = list(
           xaxis = list(title = "NMDS1"),
           yaxis = list(title = "NMDS2"),
           zaxis = list(title = "NMDS3")
         ))

html_file <- "3D_NMDS_Plot.html"
saveWidget(p, html_file, selfcontained = TRUE)

# Use webshot to take a screenshot of the HTML plot and save it as an image
image_file <- "3D_NMDS_Plot.png"
webshot(html_file, file = image_file, vwidth = 800, vheight = 600)


```

# Saves

```{r}

# PHYLOSEQ OBJECTS
saveRDS(ps, 'Scripts/Bact.Euk/Res/BE_1.PreProcessing/be_nfn.rds')
saveRDS(ps_phylo,'Scripts/Bact.Euk/Res/BE_1.PreProcessing/be_phylo.rds')

```
